Comparing the Effectiveness of the Choropleth Map with a Hexagon Tile Map for Communicating Cancer Statistics

### 2020-01-14 paper/paper.Rmd
generating trend models

<!-- Spatially dependent data sets were simulated using the variogram model on the centroids of each geographic unit. Twelve simulations from the variogram model were created for each of the twelve lineups.

The underlying spatial correlation model was created to provide spatial autocorrelation between neighboring areas using the longitude and latitude values for the Statistical Areas.

$$z = 1$$
$$locations = longitude + latitude$$
-->

@@

XXX Fill in this structure - see commented text

<!-- Discuss simulating the trend data -->

@@

The null model imposed suggests that neighbors are related. The randomness imposed was smoothed to mirror the practice taken by the Australian Cancer Atlas. This smoothing allowed neighbors to be related to each other and show distributions similar to the Liver cancer distribution shown in \ref{fig:liver}.

Spatially dependent data sets were simulated using the variogram model on the centroids of each geographic unit. Twelve simulations from the variogram model were created for each of the twelve lineups.

In these 12 sets of data, each of the 144 maps were smoothed several times to replicate the spatial autocorrelation seen in cancer data sets presented in the Australian Cancer Atlas.

For each of the 144 individual maps, the values attributed to each geographic area are rescaled to show a similar color scale from deep blue to dark red within each map.

A random location was selected for each set of lineup data.
In this location, a trend model was overlaid on the null set of spatially correlated data.
Each set of lineup data was used to produce a choropleth maps and hexagon tile maps. These matched pairs were split between Group A and Group B.

For each lineup these statistics were calculated:

 - accuracy: the proportion of subjects who detected the data plot
 - efficiency: average time taken to respond

@@

Participants were recruited to evaluate lineups from the Figure Eight crowdsourcing platform [@figeight].
The lineup protocol expects that the participants are uninvolved judges with no prior knowledge of the data, to avoid inadvertently affecting results. Potential participants needed to have achieved level 2 or level 3 from prior work. The participants were chosen as they have experience with survey tasks that involve evaluating images XXX Did we really choose participants this way??). All participants were at least 18 years old.

@@

XXX Sentence on pilot study which allowed us to estimate the effect size, and thus decide on number of participants to collect responses from

>>>>

The null model imposed by our hypothesis suggests that neighbors are related. The randomness induced when generating the null data was smoothed to mirror the practices of the Australian Cancer Atlas statisticians. 
In these 12 sets of data, each of the 144 maps were smoothed several times to replicate the spatial autocorrelation seen in cancer data sets presented in the Australian Cancer Atlas, without implementing uncertainty via transparency.

A list of neighbors for each geographic unit was generated to use when smoothing the distributions. For each geographic unit the same spatial smoother was applied in each layer of smoothing. It kept half of the units' previous value, and derived the new half as the mean of the values of its neighbors at the previous layer of smoothing.

This smoothing allowed neighbors to be related to each other, but allowed outliers, and showed distributions similar to the Liver cancer distribution shown in \ref{fig:liver}.

@@

<!-- Discuss simulating the trend data -->
The real data display for each trend type was created by applying different methods to the centroid values of each geographic unit.

The North West to Soth East (NW-SE) distribution was based on the centroid longitude and latitude values. First,  the minimum latitude and longitude value of all centroids were found. Then the difference was found between the minimums, and the centroid values of all other centroids in the set of SA3 geographic units.
The following formula was applied to the difference from the minimums: $ (-2*latitude) + longitude $.

The All Cities trend model was created using the distance from the centroid of each geographic unit to the closest capital city in Australia. For each geographic unit, the distance to the closest capital city can be found when creating the hexagon tile map using the sugarbag [@sugarbag] package.
There are large geographic differences between the centroids of greater capital city areas and the rural outback areas. 
To find the geographic units closest to the capital cities, the distance to the closest capital city was rescaled to the power of 8 to help differentiate between the smallest distances.
The distances were then scaled between 0 and 1, where very small distance from the capital city were near 0, and the farthest centroid from any capital city were near to 1. The greater capital city geographic units were considered those with a scaled distance of less than 0.85, this resulted in 201 units considered "close". This was chosen to make red spots around the cities visible, even if they were not overtly noticeable in the choropleth display.

A similar process was applied to the three cities' trend model. For each of the four replicates for the Three Cities trend, a random sample of capital cities was taken from Sydney, Brisbane, Melbourne, Adelaide, Perth, and Hobart.
The same process used for the All Cities trend was applied, but only to the areas nearest to the three cities in the sample for each of the four replicates. 
For the Three cities lineups, scaling was applied separately to the values for geographic units that were and were not closest to the cities that were included in the sample. The geographic units that were close to the three cities selected were scaled to a larger maximum value so that their greatest values would appear red.

One of the lineup map locations was randomly chosen to embed the real trend model data, in each of the four replicates, for the three trend models.
In this location, real data display of the trend model was added to the null values of the spatially correlated data for each lineup.

@@

For each of the 144 individual maps, the values attributed to each geographic area are rescaled to show a similar color scale from deep blue to dark red within each map.
This allowed at least one geographic unit to be colour dark blue, and at least one to be red, in every map display of every lineup.

For the geographic NW-SE distribution, this resulted in the smallest values of the trend model (blue) occurring in Western Australia, the North West of Australia, and the largest values of the trend model (red) occurring in the South East. This resulted in Tasmania being colored completely red.

Each set of lineup data was used to produce a choropleth maps and hexagon tile maps. These matched pairs were split between Group A and Group B according to the 2 x 3 factor experimental design depicted in \ref{fig:exp-design}.

@@

Participants were recruited to evaluate lineups from the Figure Eight crowdsourcing platform [@figeight].
The lineup protocol expects that the participants are uninvolved judges with no prior knowledge of the data, to avoid inadvertently affecting results. Potential participants needed to have achieved level 2 or level 3 from prior work. All participants were at least 18 years old.

@@

A pilot study was conducted in the working group of the Econometrics and Business Statistics Department of Monash University. This allowed us to estimate the effect size, and thus decide on number of participants to collect responses from. From the pilot study amount of participants needed to acheieve a 0.05 power level was determined to be

### 2020-01-14 paper/paper.tex
generating trend models

@@

XXX Fill in this structure - see commented text

The null model imposed suggests that neighbors are related. The randomness imposed was smoothed to mirror the practice taken by the Australian Cancer Atlas. This smoothing allowed neighbors to be related to each other and show distributions similar to the Liver cancer distribution shown in \ref{fig:liver}.

Spatially dependent data sets were simulated using the variogram model on the centroids of each geographic unit. Twelve simulations from the variogram model were created for each of the twelve lineups.

In these 12 sets of data, each of the 144 maps were smoothed several times to replicate the spatial autocorrelation seen in cancer data sets presented in the Australian Cancer Atlas.

For each of the 144 individual maps, the values attributed to each geographic area are rescaled to show a similar color scale from deep blue to dark red within each map.

A random location was selected for each set of lineup data.
In this location, a trend model was overlaid on the null set of spatially correlated data.
Each set of lineup data was used to produce a choropleth maps and hexagon tile maps. These matched pairs were split between Group A and Group B.

For each lineup these statistics were calculated:

\begin{itemize}
\tightlist
\item
  accuracy: the proportion of subjects who detected the data plot
\item
  efficiency: average time taken to respond
\end{itemize}

@@

Participants were recruited to evaluate lineups from the Figure Eight crowdsourcing platform {[}\protect\hyperlink{ref-figeight}{20}{]}.
The lineup protocol expects that the participants are uninvolved judges with no prior knowledge of the data, to avoid inadvertently affecting results. Potential participants needed to have achieved level 2 or level 3 from prior work. The participants were chosen as they have experience with survey tasks that involve evaluating images XXX Did we really choose participants this way??). All participants were at least 18 years old.

>>>>

The null model imposed by our hypothesis suggests that neighbors are related. The randomness induced when generating the null data was smoothed to mirror the practices of the Australian Cancer Atlas statisticians.
In these 12 sets of data, each of the 144 maps were smoothed several times to replicate the spatial autocorrelation seen in cancer data sets presented in the Australian Cancer Atlas, without implementing uncertainty via transparency.

A list of neighbors for each geographic unit was generated to use when smoothing the distributions. For each geographic unit the same spatial smoother was applied in each layer of smoothing. It kept half of the units' previous value, and derived the new half as the mean of the values of its neighbors at the previous layer of smoothing.

This smoothing allowed neighbors to be related to each other, but allowed outliers, and showed distributions similar to the Liver cancer distribution shown in \ref{fig:liver}.

@@

The real data display for each trend type was created by applying different methods to the centroid values of each geographic unit.

The North West to Soth East (NW-SE) distribution was based on the centroid longitude and latitude values. First, the minimum latitude and longitude value of all centroids were found. Then the difference was found between the minimums, and the centroid values of all other centroids in the set of SA3 geographic units.
The following formula was applied to the difference from the minimums: \$ (-2*latitude) + longitude \$.

The All Cities trend model was created using the distance from the centroid of each geographic unit to the closest capital city in Australia. For each geographic unit, the distance to the closest capital city can be found when creating the hexagon tile map using the sugarbag {[}\protect\hyperlink{ref-sugarbag}{1}{]} package.
There are large geographic differences between the centroids of greater capital city areas and the rural outback areas.
To find the geographic units closest to the capital cities, the distance to the closest capital city was rescaled to the power of 8 to help differentiate between the smallest distances.
The distances were then scaled between 0 and 1, where very small distance from the capital city were near 0, and the farthest centroid from any capital city were near to 1. The greater capital city geographic units were considered those with a scaled distance of less than 0.85, this resulted in 201 units considered ``close''. This was chosen to make red spots around the cities visible, even if they were not overtly noticeable in the choropleth display.

A similar process was applied to the three cities' trend model. For each of the four replicates for the Three Cities trend, a random sample of capital cities was taken from Sydney, Brisbane, Melbourne, Adelaide, Perth, and Hobart.
The same process used for the All Cities trend was applied, but only to the areas nearest to the three cities in the sample for each of the four replicates.
For the Three cities lineups, scaling was applied separately to the values for geographic units that were and were not closest to the cities that were included in the sample. The geographic units that were close to the three cities selected were scaled to a larger maximum value so that their greatest values would appear red.

One of the lineup map locations was randomly chosen to embed the real trend model data, in each of the four replicates, for the three trend models.
In this location, real data display of the trend model was added to the null values of the spatially correlated data for each lineup.

For each of the 144 individual maps, the values attributed to each geographic area are rescaled to show a similar color scale from deep blue to dark red within each map.
This allowed at least one geographic unit to be colour dark blue, and at least one to be red, in every map display of every lineup.

For the geographic NW-SE distribution, this resulted in the smallest values of the trend model (blue) occurring in Western Australia, the North West of Australia, and the largest values of the trend model (red) occurring in the South East. This resulted in Tasmania being colored completely red.

Each set of lineup data was used to produce a choropleth maps and hexagon tile maps. These matched pairs were split between Group A and Group B according to the 2 x 3 factor experimental design depicted in \ref{fig:exp-design}.

@@

Participants were recruited to evaluate lineups from the Figure Eight crowdsourcing platform {[}\protect\hyperlink{ref-figeight}{20}{]}.
The lineup protocol expects that the participants are uninvolved judges with no prior knowledge of the data, to avoid inadvertently affecting results. Potential participants needed to have achieved level 2 or level 3 from prior work. All participants were at least 18 years old.
